---
title: "Pins Demo"
author: "Jeremy Allen"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
  chunk_output_type: console
---

# Pins

The pins package helps you publish data sets, models, and other R
objects, making it easy to share them across projects and with your
colleagues. The pin board can be on

-   local or temp folder
-   Shared folder (like Dropbox)
-   RStudio Connect
-   Amazon S3
-   Azure storage
-   Microsoft 365 (OneDrive and SharePoint) - Google Cloud

![flow diagram for using Pins with RStudio
Connect](example_pins_workflow.png)

## I don't quite get the use case?

A "yes" to any of these indicates a good use case for pins!

-   Using read_csv() to bring in emailed data?
-   Saving .Rds or .RData objects to be called later?
-   Sharing your model or data across multiple apps?
-   Redeploying your app every time the supporting data is updated?

Now revisit that flow diagram then consider the examples below.

```{r, eval=F, echo=T}

library(pins)
library(fs)

```

What version of pins are we using?

```{r, eval=F, echo=T}

packageVersion("pins")

```

## AWS S3 buckets as pin boards

Note the use of environment variables for my key ID and key.

#### Create the board

```{r, eval=F, echo=T}

# Amazon AWS S3 board

bs3 <- board_s3(
  Sys.getenv("AWS_BUCKET"),
  prefix = "demo",
  versioned = TRUE,
  access_key = Sys.getenv("AWS_ACCESS_KEY_ID"),
  secret_access_key = Sys.getenv("AWS_SECRET_ACESS_KEY"),
  session_token = NULL,
  credential_expiration = NULL,
  profile = NULL,
  region = Sys.getenv("AWS_REGION"),
  endpoint = NULL,
  cache = NULL
)

```

#### What's already on the board in my S3 bucket

```{r, eval=F, echo=T}

# anything on there now?
bs3 %>% pin_list()

```

#### What data sets come with R?

```{r, eval=F, echo=T}

data()

```

#### Write more to the board in my S3 bucket

```{r, eval=F, echo=T}

# write new data frame to the board in my S3 bucket
bs3 %>% pin_write(Indometh) #insert name of data frame

```

pins wrote the data frame as an rds file

#### Pause. Let's do the same for a local board

creates a board in a system data directory. It's useful if you want to
share pins between R sessions on your computer, and you don't care where
the data lives.

```{r, eval=F, echo=T}

my_local_board <- board_local()
my_local_board %>% pin_list()
my_local_board %>% pin_delete("attitude")
my_local_board %>% pin_list()
my_local_board %>% pin_write(attitude) 
my_local_board %>% pin_list()

```

#### And let's do the same for a folder board

creates a board inside a folder. You can use this to share files by
using a folder on a shared network drive or inside a DropBox.

```{r, eval=F, echo=T}

# if folder is not there it will be created
my_proj_board <- board_folder(path = "project_board")

my_proj_board %>% pin_list()
my_proj_board %>% pin_delete("attitude")
my_proj_board %>% pin_write(attitude) 
my_proj_board %>% pin_list()

```

#### Let's look at the board in our project directory

There's a .rds file and a text file. The text file contains metadata
about the pinned object

#### Interact with the pinned objects on boards

```{r, eval=F, echo=T}

bs3 %>% pin_list()

bs3 %>% pin_read("islands")

bs3 %>% pin_search(search = "CO2")

bs3 %>% pin_delete("CO2")

my_proj_board %>% pin_delete("attitude") 

# add metadata to a pin
my_proj_board %>% pin_write(attitude,
                            "attitude",
                            versioned = TRUE,
                            description = "2007 data about bad cat attitudes", 
                            metadata = list("year" = 2007,
                                            "checked_by" = "Jeremy"))

my_proj_board %>% pin_meta("attitude") 
my_proj_board %>% pin_versions("attitude") 

```

## RStudio Connect as a board

```{r, eval=F, echo=T}

# RStudio Connect board

# make sure package rsconnect is already installed
# Connect boards are versioned by default
brsc <- board_rsconnect() 
brsc %>% pin_list() # grab a coffee while this runs

brsc %>% pin_search(search = "codenames")

cn <- brsc %>% pin_read(name = "katie.masiello/codenames")

View(cn)

```

## Versioning

board level or pin level

```{r, eval=F, echo=T}

# all pins on this board will be versioned
my_version_board <- board_folder(path = "project_v_board",
                              versioned = TRUE)

my_version_board %>% pin_write(Seatbelts,
                            versioned = TRUE)

my_version_board %>% pin_list()

my_version_board %>% pin_versions("Seatbelts")

df_versioned <- my_version_board %>%
  pin_read(
    name = "Seatbelts",
    version = "20220914T141335Z-08fde"
  )

```

## Pin Models

See [Bike Predict example
project](https://solutions.rstudio.com/example/bike_predict/) to see how the vetiver package is used to pin a model to Posit Connect.  

## Upload a file instead of pinning

This is a lower-level interface than pin_read() and pin_write() that you
can use to pin any file, as opposed to any R object. The path returned
by pin_download() is a read-only path to a cached file: you should never
attempt to modify this file.

```{r, eval=F, echo=T}

# upload file instead of pinning R objects

my_proj_board %>% pin_upload(
  paths = "msg.txt",
  name = "message_in_bottle",
  title = "Shared Bottle Message",
  description = NULL,
  metadata = NULL
)


my_proj_board %>% pin_list()

# let's go to a new session and download that text file

```

## How to get started

Katie's [great PDF
guide](https://colorado.rstudio.com/rsc/protip-pins/protip_pins.pdf "Katie's getting started guide!")
that lives on our public Connect.

Katie's [great video demo](https://www.youtube.com/watch?v=3Yk6U_XqxNo),
including versioning pins and models

Our [Bike Prediction example
project](https://solutions.rstudio.com/example/bike_predict/) with free
code!
