---
title: "Road Less Traveled `r emo::ji('national park')` `r emo::ji('person mountain biking')`"
subtitle: "Productive RStudio IDE features you may not know"
format: html
editor_options: 
  chunk_output_type: console
bibliography: references.bib
---

...and some you will

```{r}
#| echo: true
#| warning: false
#| error: false

# to install the emo package, first install the devtools package,
# then install emo from github with: devtools::install_github("hadley/emo")

library(emo)
library(tidyverse)

```

<br>

::: callout-note
I am on a MacBook so `cmd` = command key. If you are on a Windows machine, use `ctl` instead
:::

<br>

## `r emo::ji("red heart")` Jeremy's personal favorites

### Output to where?

I like it to go to the console

-   `r emo::ji("gear")` -\> *Chunk Output in Console*

### Console is where?

I move my console to the top right quadrant: *Tools*-\>*Global Options*-\>*Pane Layout*

### Function bodies

-   F2 or `cmd+click` the function name to expose function body\
-   Don't forget to select a method in the drop down if you can
-   F1 the function name for help
-   My favs because I learn best by seeing what others are already doing `r emo::ji("brain")`

```{r}
#| echo: true
#| eval: true
#| warning: false
#| error: false

emo::ji("red heart")

dat <- tibble::as_tibble(diamonds) 

```

### View the data in a tab

-   `cmd+click` the tibble name
-   search whole table
-   filter columns

<br>

## `r emo::ji("compass")` Getting around

### Command Palette

It's not unknown, but it's so productive I cannot not mention it

-   `shift+cmd+p`
-   once in there, type Global Options and return
-   explore Global Options
    -   keybindings, vim anyone?
    -   relative line numbers
    -   highlight function calls (toggle in Command Palette)

### Outline is not out of line

-   `ctl+shift+o` to navigate your document with the outline on the right
-   Also access outline and functions at the bottom of the doc

### Death by a thousand tabs

-   *Go to file/function window* - find function grow_city()
-   `ctl+.` to search file tabs

### It's ok to cheat - I don't believe they've been here all along

-   *Help* -\> *Cheat Sheets*
-   What? There's one for the IDE? `r emo::ji("face with monocle")`

<br>

## `r emo::ji("computer")` Code mode

### Editor fun

-   `shift+ctl+c` to comment all highlighted lines
-   `alt+up` or `alt+down` to move lines
-   `ctl+alt+down` or `ctl+alt+up` or `alt+drag` to have multi-line cursor
-   `cmd+i` to fix indention in highlighted code
-   `cmd+shift+m` to insert pipe, but which pipe? See `shift+cmd+p`
-   in the console, what does `up` do? what about `cmd+up`?
-   double-click parentheses to highlight contents

```{r}
#| echo: true
#| eval: true
#| warning: false
#| error: false

dat |>
  ggplot2::ggplot(aes(x = carat, y = price, color = cut)) +
  geom_point(alpha = .3) +
  labs(
    title = "Diamonds: Price as a Function of Carat",
    subtitle = "by cut",
    x = NULL,
    y = NULL
  ) +
  theme_minimal()

```

### Snippets

-   `shift+tab` to expand a snippet
-   type fun then `shift+tab`
-   don't stop now, do `if` and `for` as well
-   shinyapp then `shift+tab`

```{r}
#| echo: true
#| eval: false
#| warning: false
#| error: false

fun

if

for

shinyapp

```

-   *Tools* -\> *Edit Code Snippets* to see them all and make your own

<br>

## `r emo::ji("hospital")` Search and Rescue

-   find and replace across files
-   `shift+cmd+f` and search for good_pets, then replace
-   but wait, we didn't search .qmd files, so use custom filter

<br>

## `r emo::ji("chains")` Connections

### The IDE installs Posit's pro drivers for you

In the Connections pane

-   click *New Connection*
-   chose a driver
-   the IDE installs it before helping you start the connection
-   if the install fails, update packages and try again `r emo::ji("plug")`

<br>

## `r emo::ji("beetle")` Bugs

### Breakpoints and browser()

Let's jump over to the `my_func.R` file for this piece

-   step through the body of a function
-   see variable values along the way
-   choose which environment to see

<br>

## `r emo::ji("button")` Visual editor

-   normal doc styling features

-   insert links, pics, tables, citations, footnotes, and more

-   Here are two citations [@Little2020; @Melnychuk2017]; and a reference section will be added to the doc

-   add more here: [@Welcomme2010; @Teh2017]

<br>

## `r emo::ji("package")` Advanced

Require writing packages

### Addins

[RStudio Addins](https://docs.posit.co/ide/user/ide/guide/productivity/add-ins.html)

### Templates

[Project Teamplates](https://docs.posit.co/ide/user/ide/guide/productivity/project-templates.html)

<br> <br>

## So Much More

### Have you tried background jobs yet?

Tip: you can activate one from inside an R file - *Source* -\> *Source as Background Job*

[RStudio User Guide!](https://docs.posit.co/ide/user/)
